Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-04-03 conclusion.tex
Iteration conclusions

The PanDA  system was developed to meet the scale and complexity of LHC
distributed computing for the ATLAS experiment.  In the process,  the old batch
job paradigm  of computing in HEP was  discarded  in favor of a  far more
flexible and scalable  model. The success  of PanDA  at the LHC is leading to
widespread adoption and testing by other experiments. PanDA  is the first
exascale  workload management system in HEP, already operating at a million
computing jobs per day, and processing over an exabyte of data in 2013. Next LHC
run will pose massive computing  challenges. With a  doubling of the beam
energy  and luminosity as  well as an increased  need for  simulates  data, the
data volume is expected to increase with a factor 5--6 or more. Storing and
processing  this amount of data is a  challenge   that cannot be resolved with
the currently existing  computing  resources in ATLAS\@. To resolve this
challenge, ATLAS is turning to commercial  as well as academic Cloud services
and HPCs via the PanDA system. Also the work underway is enabling the use of
PanDA by new scientific collaborations and communities as a means  of leveraging
extreme scale computing  resources with a low barrier of entry. The technology
base provided by the PanDA system will enhance the usage of a variety  of
high-performance computing resources available to basic research.

HEP was amongst the first, if not the very first experimental community to
realize the importance of using WMS to manage their computational campaign(s).
As computing become increasingly critical for a range of other experiments,
the impact of WMS on HEP foreshadows the importance of WMS for other
experiments and their computational campaigns. Other experiments such as SKA,
LSST etc.., will have their own workload characteristics, resources types and
federation constraints, as well metrics of performance.  The design and
experience captured in this paper should proof invaluable for computational
campaigns that need to evaluate the relative merits of different approaches.

Thanks to the efforts captured in this paper, the time-averaged utilization of
Titan has consistently been upwards of 90\%. Compare this to the time averaged
utilization of other similar leadership machines such as NSF's flagship Blue
Waters, where the average utilization fluctuates between 60-80\% (XDMoD)
machine~\cite{bw-sucks}. These impressive if not unprecedented operation gains
aside, this work is just a starting point towards more effective operational
models for future leadership and online analytical platforms\cite{foap-url},
which will have to support ever increasing complex workloads with varying
models for dynamic resource federation. Recent efforts with Google Cloud for
the CMS experiment -- HEPCloud~\cite{hepcloud,google-hep} represent similar
yet different approaches to dynamic resource federation.

Our experience demonstrated the advantages arising from a  systems engineering
and abstractions based approach, in particular the pilot abstraction. As the
type, heterogeneity and complexity of future platforms  (e.g., Exascale,
Commercial Clouds) for science increases, there will be a greater need and
emphasis on abstractions to ensure WMS can both utilize as well as support
platforms.

>>>>

% PanDA WMS was developed to meet the scale and complexity of LHC
% distributed computing for the ATLAS experiment. In the process, the old batch
% job paradigm of computing in HEP was discarded in favor of a far more
% flexible and scalable model. The success of PanDA at the LHC is leading to
% widespread adoption and testing by other experiments. PanDA is the first
% exascale workload management system in HEP, already operating at a million
% computing jobs per day, and processing over an exabyte of data in 2013. Next
% LHC run will pose massive computing challenges. With a doubling of the beam
% energy and luminosity as well as an increased need for simulates data, the
% data volume is expected to increase with a factor 5--6 or more. Storing and
% processing this amount of data is a challenge that cannot be resolved with the
% currently existing computing resources in ATLAS\@. To resolve this challenge,
% ATLAS is turning to commercial as well as academic Cloud services and HPCs via
% the PanDA system. Also the work underway is enabling the use of PanDA by new
% scientific collaborations and communities as a means of leveraging extreme
% scale computing resources with a low barrier of entry. The technology base
% provided by the PanDA system will enhance the usage of a variety of
% high-performance computing resources available to basic research.

The deployment of PanDA Broker on Titan enabled distributed computing on a
leadership-class HPC machine at unprecedented scale. In the past 13 months,
PanDA WMS has consumed almost 52M core-hours on Titan, simulating 3.5\% of the
total number of detector events of the ATLAS production Monte Carlo workflow. We
described the implementation and execution process of PanDA WMS
(\S\ref{sec:panda_overview}) and PanDA Broker (\S\ref{sec:panda_titan}), showing
how they support and enable distributed computing at this scale on Titan, a
leadership-class HPC machine managed by OCLF at ORNL.

We characterized the state of practice by evaluating the efficiency, scalability
and reliability of both PanDA Broker and AthenaMP as deployed on Titan
(\S\ref{sec:panda_titan}). Our characterization highlighted both the strengths
and limitations of the current design and implementation: PanDA Brokers enable
the sustained execution of millions of simulations per week but further work is
required to optimize its efficiency and reliability (\S\ref{ssec:broker_titan}).
PanDA Brokers support the concurrent execution of multiple AthenaMP instances,
enabling each AthenaMP to perform the concurrent execution of up to 16 Geant4
simulators. Nonetheless, our characterization showed how improving I/O
performance could reduce overheads (\S\ref{ssec:athenamp_titan}), increasing the
overall utilization of Titan's backfill availability.

More in general, this paper highlighted the fundamental role of WMS for
experimental science. HEP was amongst the first, if not the very first
experimental community to realize the importance of using WMS to manage their
computational campaign(s). As computing become increasingly critical for a range
of other experiments, the impact of WMS on HEP foreshadows the importance of WMS
for other communities.
% experiments and their computational campaigns.
Other experiments such as SKA or LSST, will have their own workload
characteristics, resources types and federation constraints, as well metrics of
performance. The design and experience captured in this paper should prove
invaluable for computational campaigns that need to evaluate the relative merits
of different approaches.

Thanks to the efforts captured in this paper, the time-averaged utilization of
Titan has consistently been upwards of 90\%. Compare this to the time averaged
utilization of other similar leadership machines such as NSF's flagship Blue
Waters, where the average utilization fluctuates between 60-80\% (XDMoD)
machine~\cite{bw-sucks}. These % impressive if not
unprecedented operation gains aside, this work is just a starting point towards
more effective operational models for future leadership and online analytical
platforms\cite{foap-url}, which will have to support ever increasing complex
workloads with varying models for dynamic resource federation. Recent efforts
with Google Cloud for the CMS experiment -- HEPCloud~\cite{hepcloud,google-hep}
represent similar yet different approaches to dynamic resource federation.

Our experience demonstrated also the advantages arising from a systems
engineering and abstractions based approach, in particular the pilot
abstraction. As the type, heterogeneity and complexity of future platforms
(e.g., Exascale, Commercial Clouds) for science increases, there will be a
greater need and emphasis on abstractions to ensure WMS can both utilize as well
as support platforms.
